---
title: "Hurricane Effects on Stream Chemistry"
format: html
---

# Background

A study (Schaefer et al. 2000) investigated the effects of stream water quality in Puerto Rico after the 1989 Hurricane Hugo. They analyzed the level of nutrients over time and found that potassium levels stayed elevated for years afterwards.

## Data
Load relevant libraries and source data. 
```{r}
library(here)
library(tidyverse)
library(janitor)
library(ggplot2)

source(here("R", "1_clean_data.R"))
source(here("R", "rolling_mean_function.R"))
source(here("R", "ggplot.R"))
```
Import data 
```{r}
bq1 <- read_csv(here::here("data", 
                           "stream-water-pr",
                           "QuebradaCuenca1-Bisley.csv")) %>% 
  clean_names() 

bq2 <- read_csv(here::here("data", 
                           "stream-water-pr",
                           "QuebradaCuenca2-Bisley.csv")) %>% 
  clean_names()


bq3 <- read_csv(here::here("data", 
                           "stream-water-pr",
                           "QuebradaCuenca3-Bisley.csv")) %>% 
  clean_names() 


mpr <- read_csv(here::here("data", 
                           "stream-water-pr",
                           "RioMameyesPuenteRoto.csv")) %>% 
  clean_names() 
```
# Methods

## 1. Clean raw data files
The data cleaning process was as follows:
1. Loading in raw data. 
2. Clean names: changing the names from upper case to lower-snake-case. 
3. Filter data by only relevant chemicals/nutrients in the study.
4. 


## 2. Apply rolling mean function. 

```{r}
source(here("R", "1_clean_data.R"))
streams <- read_csv(here("processed-data", "streams.csv"))

#View(streams_longer) #use KablEextra 

```

## Results
